|

### 2021-11-14 07-Chp-Triangulation.Rmd
End of day:

Triangulation is the practice of using multiple sources of evidence to provide more reliable answers to a research question. Different sources of evidence will have different limitations. If the result from each source points towards the same answer, this improves our confidence in the conclusion.

@@

## Introduction

This chapter will attempt to triangulate the evidence identified by the systematic review in Chapter \@ref(sys-rev-heading-results), with the the results from the analysis of CPRD data (Chapter \@ref(cprd-heading) and IPD \@ref(ipd-heading).

@@

### Overview of triangulation {#triangulation-overview}

Aetiological triangulation, or simply triangulation, is the process of comparing and contrasting across result in order to .

It has been widely used in the literature, and is comparable to Bradford-Hill's criteria of "consistency", that is the replication of an observed relationship across several different contexts.[@hill1965] A good working definition of the triangulation is:

@@

This represents a significant step forward from the current evidence synthesis practice of synthesising evidence from only one approach (e.g. RCT, observational study, etc.).

However, previous attempts at triangulation have ignored the potential for varying and opposing sources of bias within an approach, assigning a level and direction of bias to entire categories of studies.

Recent development in bias assessment have a allowed fro a much more granular consideration of the threats to the internal validity of a result.

For example, while it may be true that all non-randomised studies share some _minimal level_ of bias, investigation of the biases applicable to each individual result gives useful additional information, and is essential for any attempt at quantitative triangulation.

This chapter, in addition to providing a narrative comparison of the existing and new evidence identified in this thesis, attempts to combine the results numerically by building on existing work on this topic.

In the following section, I define important terms used throughout this chapter.

@@

__Different types of triangulation__

Triangulation exercises can be characterised as either qualitative or quantitative.

_Qualitative_ triangulation is describing the different results. It regularly appears though not under that name, in the discussion of results section under the name "comparison with other literature".

In contrast, _quantitative triangulation_ attempts to obtain a overall numerical result by synthesising the different types of evidence. This method is in its infancy, given important methodological challenges that are illustrated in this Chapter.

@@

As outlined in the thesis overview presented in Section \@ref(background-thesis-overview), this triangulation exercise draws on the research produced in each of the preceding chapters. More specifically, this chapter builds on the comprehensive systematic review presented in chapter \@re(sys-rev-heading), and incorporates the results of the analyses presented in Chapters \@ref(cprd-heading) (assocation of statins with dementia outcomes in the CPRD) & \@ref(ipd-heading) (association of lipids with dementia outcomes in IPD).

Table \@ref(tab:thesisOverview-table) illustrates the causal question each approach has attempted to address, along with the exposures and outcomes for each.

\blandscape
<!----------------------------------------------------------------------------->
(ref:thesisOverview-caption) Summary of studies included in this thesis, and used as evidence sources in the triangulation exercise. Note, Chapter \@ref(sys-rev-tools-heading) is intentionally not included in this table, as it describes a tool rather than a research study.

@@

As a first step to triangulating the available evidence, all information sources were grouped by outcome, and systematically compared and contrasted. Potential reasons for heterogeneity between study designs were examined with specific reference to the risk-of-bias assessments performed.

@@

In addition to the qualitative discussion of evidence, I attempted to integrate the numerical results of the multiple approaches. This proof-of-concept approach incorporated advancements in the way that bias in results is assessed (domain-based assessment tools) and existing methods for bias-adjusted meta-analysis[@turner2009] to to illustrate how causal questions could be addressed under this quantitative triangulation framework.

Several steps were used to accomplish this approach, as described in the following sections. In summary these steps are:

1. Define the causal question of interest
1. Identify of relevant evidence sources
1. Assess(and visualise) the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Perform bias-adjusted meta-analysis using bias-adjusted results

The approach was implemented for two causal questions of interest, as defined below.

@@

<!-- TODO See ROBINS-E for this, as to how to think about exposures, adn the "Duration and timing of exposure being assessed with different approachesDuration and timing of exposure being assessed with different approaches" section of the aetiological triangulation -->

This was also guided by the forthcoming ROBINS-E tool, which has

*	Effect of average exposure to LDL-c during midlife on AD 
* Effect of late-life LDL-c on all-cause dementia

&nbsp;

@@

Once the causal question of interest had been defined, studies related to it were

individual results were obtained from the general evidence sources described in Section \@ref(tri-data-sources)

In order to illustrate the bias-/indirectness adjustment process detailed in the subsequent sections, the process for calculating the adjusted estimate for a single study is described in detail for a single study. Following this, the process was applied to the two subsets of studies related to the causal questions of interest.

@@

The risk of bias tools used to assess each result are described in detail in Section \@ref(risk-of-bias). In summary, randomized controlled trials were assessed using the RoB 2 tool;[@sterne2019] non-randomised studies of interventions (NRSI), were assessed using the ROBINS-I tool;[@sterne2016]

For the sake of consistency, I mapped the different acceptable judgements in each risk-of-bias tool to a harmonised set of judgements: "Low", "Moderate","High". The exact mapping can be seen in Table \@ref(tab:robLevelsMapping-table). Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis. <!-- TODO CITATION NEEDED -->

@@

In addition to assessing the extent of bias in each domain, as defined by the three levels detailed above, I attempted to records the direction of the bias, so that it could feed into the triangulation.

This was achieved via an additional question at the end of each bias domain, which asked assessors to predict the expected direction of bias in that domain. Note that this question is only apply to existing tools (ROB2/ROBINS-I/ROBINS-E), but I employed an identical approach during assessment of Mendelian randomisation studies. This optional question response were:

* "Favours experimental"/"Favours comparator" (additive bias)
* "Towards null"/"Away from null" (proportional bias)
* "Unpredictable"

As indicated in the options above, the response t this prompt is used both to define the direction and type (additive or proportional) of bias (see ).

When predicting the direction of effect

From these response the direction of effect was calculated. For additive biases, this was simply a case of shifting the effect estimate to the left or right. For proportional biases, decide

For example, the direction of the bias when the direction of bias is thought to favour the intervention/exposure will be be same regardless of the position of the point estimate. The same is not true for the "Towards the null"/"Away from the null" options.

For example, if the effect estimate represents a protective effect (below the null), then bias towards the null would be represented by a right pointing arrow (Figure \@ref(fig:)). In contrast, if the effect of the intervention is harmful (effect estimate above the null), bias towards the null will be represented by a right-pointing arrow. The program to create these figures thus requires knowledge of both the predicted direction of bias and the effect estimate.

However,

S

Highlight that this is slightly different for the confounding domain in non-randomised studies, which only allows for the "Favours experimental", "Favours comparator", and "Unpredictable" options.

Of note, the guidance for the tools suggests that if assessors are not able to articulate why the bias in a given domain will "pull" the observed estimate in a specific direction, they should not try to guess.

To assess the usefulness of this system, the number of domains for which bias was expected (i.e. judged to be a more than Low risk of bias) but for which no expected direction could be identified were expressed as a percentage. If this percentage is high, the helpfulness of the approach is

<!-- TODO text -->

@@

To aid with the triangulation exercise and to act as, I created a new method of presentation of the extent/direction bias asessment results to enable detailed comparison across different studies contributing to the synthesis. While these figures do not

Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), the level of bias across the domains in each study is reported using coloured blocks. However, in contrast to the previous plots, arrows are used in place to illustrate the predicted direction of bias in that domain, categorised as pulling towards the left or the right.

This second aspect required some additional consideration, given the distinct and competing options available to assessors.

Tha

Helper functions in the forthcoming `triangulate` package - see Appendix \@ref() - help to preprocess risk of bias data for use with both bias-direction plots and adjusted meta-analyses. These bias direction plots themselves were created using the risk of bias isualisation tool described in Appendix \ref().

<!-- TODO include illustrative example here -->

Where the direction of bias was unclear/could not be determined from the report, this is indicated by a dash.

<!-- TODO Does immortal time bias not also have a consistent effect? yes, but inception bias does not necessarily -->

@@

####	Assign modifying values to risk/direction of bias

<!-- TODO Note that they are on the log scale - Lines 45-50 of addbias -->

For example, for a "High" additive bias in $j$th domain of the $i$th study is deifined as $\delta_{i,j}^{\mathrm{High}}$ and uncertainty around this estimate is given by a distribution

@@

The sign of $\mu_{i,j}^{\mathrm{High}}$ is defined by the relative direction of the bias in relation to the point estimate. If the bias is expected to pull the effect to the left, the position is given a negative sign, and if it is expected to pull the effect to the right, it is given a positive sign.

One key feature of the domain-based risk-of-bias domains is that the domains are considered interchangeable - i.e. a "high" bias in one domain is equivalent to a "high" bias in any other. As such, for this analysis, I assumed that the distributions assigned to each level of bias are identical across all values of $j$, that is, are identical for all domains in the risk of bias assessment tool.

@@

@@

(ref:priorsAdd-caption) Distributions for additive bias assigned to each bias judgement under two sensitivity scenarios, on the log scale.

@@

For the purposes of this illustrative analysis and to highlight the generalised nature of the code used to perform the analysis (see Section \@ref(tri-software)), I defined two sets of modifying distributions (Table 1). In the first scenario, the relationship between the position assigned to each judgement is linear. In the second scenario, . In both cases, the variance of the "High" risk of bias judgement was defined to be great than the "Moderate" risk of bias judgement. The two scenarios are illustrated in Figure \@ref(fig).

Reasonable values for the position and spread of the distributions were informed by calculating the mean and maximum adjustment values as defined by the original elicitation exercise (see Appendix \@ref(?) for more details on how these were calculated). [@turner2009] <!-- TODO Link this to open data! -->

Where the direction of bias in a domain was unpredictable for a given result, the mean of the adjustment distribution for that domain, but the appropriate variance for (i.e. no predicted direction, but increased uncretainty in line with judgement).

@@

A similar approach was used to define prior distributions for external bias or indirectness. I used a similar enforced scale of "Low/"Moderate"/"High" indirectness versus each of the 4 aspects of the target causal question. (recap).

In the original method, this was assessed against 4 categories of external bias (population/intervention/control/outcome)

In this

A key example here is that

All indirectness bias were defined as being proportional. As described i

This is also in line with previous efforts to assess the adjustment needed due to indirectness

As crude example focusing on the intervention , versus the causal question of the effect of LDL-c in mid-life on Alzheimer's disease, studies of lipids levels in this time period would require minial adjustment, whereas studies
...

@@

Once the total additive and proportional biases for each result were computed, they were used to define an adjusted estimate for each result included in the synthesis.

@@

where $\mu_{i\delta}$ denotes total additive and $\mu_{i\beta}$ the total proportional adjustment, and the $^{\mathrm{B}}$ and $^{\mathrm{In}}$ superscripts denote wether the adjustment relates to internal bias or indirectness respectively.

Using the same terms, the standard error of this estimate, $\hat{\theta}$, is then calculated as:

@@

Once each individual result has been adjusted, they are combined n a standard random-effects meta-analysis <!-- TODO CITATION NEEDED -->

#### Software {#tri-software}

Personal communication with the authors of the original method resulted in the STATA code to implement the

As detailed in the introduction to this section, the model for defining bias-indirectness-adjusted results was available as STATA code. The  has since been generalised as part of the `triangulate` R package to allow for users to apply the approach detailed here.

The package allows for preprocessing of domain-based risk-of-bias data to correctly identify the direction of the bias relative to the effect estimate, calculate bias/indirectness-adjusted estimates for each result, and synthesise these adjusted results in a standard random-effects meta-analysis.

All meta-analyses conducted within the package are implemented using the `metafor` R package.

@@

@@

Calculate both-adjusted

@@

#### Causal question #1

Midlife LDL-c on Alzheimer's disease

@@

(ref:ldlAdBiasDirection-cap) __Bias direction plot, summarising internal biases related to effect of LDL-c on Alzheimer's disease :__

@@

```{r ldlAdBiasDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:ldlAdBiasDirection-cap)', out.width='100%', fig.scap='(ref:ldlAdBiasDirection-scap)'}
knitr::include_graphics(file.path("figures/tri/midlife_AD.png"))
```
&nbsp;<!----------------------------------------------------------------------->

Including sensitivity analyses

and citations of studies - going to be a bit of pain

As expected

<!----------------------------------------------------------------------------->
(ref:fpLdlAd-cap) __shortcap:__

(ref:fpLdlAd-scap) shortcap

```{r fpLdlAd, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:fpLdlAd-cap)', out.width='100%', fig.scap='(ref:fpLdlAd-scap)'}
knitr::include_graphics(file.path("figures/tri/fp_paired_midlife_ldl_ad.png"))
```
<!----------------------------------------------------------------------------->

@@

#### Causal question #2

<!----------------------------------------------------------------------------->
(ref:ldlDemBiasDirection-cap) __Effect of LDL-c on All-cause dementia in late life:__

(ref:ldlDemBiasDirection-scap) Effect of LDL-c on All-cause dementia in late life

```{r ldlDemBiasDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:ldlDemBiasDirection-cap)', out.width='100%', fig.scap='(ref:ldlDemBiasDirection-scap)'}
knitr::include_graphics(file.path("figures/misc/UoBcrest.pdf"))
```
<!----------------------------------------------------------------------------->

@@

Summary of risk-of-bias/triangulation results

Percentage of domains for which a direction of bias could be assigned?

Main takeaway is that heterogeneity between study designs is substantially reduced

Compare bias-adjusted and

See what happens if using additive (l=1,m=2,s=3,c=4) vs multiplicative/log scale (l=1,m=2,s=4,c=8). Present adjusted results and meta-analysis of adjusted results.

Applicability/indirectness as an issue 
Compare with GRADE and cite George's example of 40% of effect predicted by MR seen when using statins for 5 years

@@

Thsi chapter has narratiely synthesised the exsiting and new evidence from the previous chapters, highlighting the lim

In addition, it has sought to provide a generalised framework for quantitative triangulation, building on existing methods for systematic domain-based risk of bias assessment and bias-/indirectness-adjusted meta-analysis.

Future quantitative triangulation should move beyond the concept of comparing results at the approach level, and instead focus on the inherent threats to internal validity present in each specific result relevant to the causal question. While this makes the process both more labour intensive and complex, to assume that, for example, all retrospective studies are equally at risk is to overlook.

@@

###	Limitations and strengths

@@

#### Problems with quantitative triangulation approach

Missing natural variation obtained from elicitation approach. Trade-off between systematic approach presented here and personal variation obtained from several reviewers.

A potential combination between the two would be to use the variation in risk-of-bias assessments

No good information on what the potential distributions for different levels of bias (High vs Moderate vs Low) should be.

@@

For example, if it was . New software for performing risk of bias assessments, being developed by the Bristol Appraisal and Review of Research group has this as a secondary methodological aim.

In addition, the generalised framework will allow for

Appreciation that both the risk of bias and indirectness of a result should be assessed is growing. Indeed, some existing risk of bias for the assessment of diagnostic test accurary[@whiting2011a] and prediction models[@moons2019] consider indirectness (termed applicability in these tools) alongside the assessment of bias. <!-- TODO Cite PROBAS and QUADAS -->

#### Subjectivity of risk of bias assessments

By definition, . In this case of this proof-of-concept analysis, the direction of bias was assessed by a single assessor. If attempting to impact policy, the

#### Defining appropriate values for bias/indirectness adjustment

@@

In the mean time, the extent of the impact of knowledge of the results could be analysed using existing data. For example, a re-analysis of the data presented in a recent paper where both expert eliciation and structured

@@

Similar there are several examples of usmeta to define priors for other statistical terms in meta-analysis. For example, previous work has looked at het

Like other studies examining the prior distributions of heterogeneity across different topic areas <!-- TODO Cite turner etc -->

Talk about different ways to define bias terms, using expert-elicitation vs .[@rhodes2020]

Also, talk about the fact that much of the bias-adjusted analysis work to date has been done on trials. Describe reasons for this but highlight that for triangulation, there needs to be better coverage of

@@

#### Exclusion of studies at critical risk of bias

@@

#### Type of bias (proportional vs additive) not shown on bias direction plot

In contrast

The graphics shown in Section \@ref() do not show the type of bias (additive or proportional), only the direction. While this is acceptable, future work could assess, potentially by using different arrow types to denote the different types of bias.

A key strength of this project is the development of a standardised way of capturing bias domains and converting the.

Helper functions in the triangulate package convert risk-of-bias assessment

@@

#### Challenges of real-life data

@@

#### Standardisation across study designs

Standardising across different measures of ex

>>>>

__TBC__

<!-- Triangulation is the practice of using multiple sources of evidence to provide more reliable answers to a research question. Different sources of evidence will have different limitations. If the result from each source points towards the same answer, this improves our confidence in the conclusion. -->

@@

@@

## Introduction {#triangulation-overview}

Aetiological triangulation, or simply triangulation, is the process of comparing and contrasting across different sources of evidence. Triangulation is broadly comparable to Bradford-Hill's criteria of "consistency", that is the replication of an observed relationship across several different contexts,[@hill1965] where contexts are assumed to have different underlying bias structures. More formally it can be defined as:

@@

This approach represents a significant step forward from the current evidence synthesis practice of synthesising evidence from only one approach (e.g. randomised controlled trials (RCTs), non-randomised studies of exposures (NRSE) or interventions (NRSI), etc.). The most common implementation of this method to date has been in the form of _qualitative triangulation_ - that is the identification and a narrative comparison, where the internal biases and differences between the study question and the causal question of interest (indirectness) of each result are discussed.

However, qualitative triangulation faces issues at scale. There is a need to include all evidence, as otherwise the potential for cherry-picking, yet as the number  as it is not possible to compare and contrast across multiple results in any meaningful way. This fact is also illustrated by previous exemplars of triangulation only considering three individual results. <!-- TODO CITATION NEEDED - original triangulation paper and Ference --> An alternative option previously employed is to assign bias to the output of a meta-analysis of similarly designed studies, but this loses information on the specific biases inherent to each results contributing to the summary estimate. For example, while it may be true that all non-randomised studies share some minimal level of bias, investigation of the biases applicable to each individual result gives useful additional information. Recent development in bias assessment have allowed for a much more granular consideration of the threats to the internal validity of a result. In combination, these limitations to qualitative triangulation detail a need for a more systematic way to integrate across multiple evidence sources as the number of individual results contributing to the exercise increases.

This chapter, in addition to providing a narrative comparison of the existing and new evidence identified in this thesis, builds on recent developments in risk-of-bias assessment and existing methods for bias-/indirectness-adjusted meta-analysis to propose a generalised framework for quantitative triangulation. The method, along with the methodological challenges to its implementation, is illustrated via a proof-of-concept case study examining the causal effect of average exposure to LDL-c at mid-life on Alzheimer's disease risk.  In the following section, I define important terms used throughout this chapter.

@@

For the sake of clarity, and given the multiple competiting defitions used to define key concepts on this topic, I begin by defining the terms as used in this chapter.

<!-- TODO Does Julian have a good reference for the conflict between different terms used for biases? -->

__Bias vs indirectness__

@@

This triangulation exercise draws on the research produced in each of the preceding chapters. More specifically, this chapter builds on the comprehensive systematic review presented in chapter \@ref(sys-rev-heading). It also incorporates the new evidence produced by the analysis of the association of statin use with dementia outcomes in the CPRD (Chapter \@ref(cprd-heading)) and of the association of lipids with dementia outcomes in previously unanalysed datasets accesssed via the DPUK (Chapter \@ref(ipd-heading)).

Table \@ref(tab:thesisOverview-table) summarises the research each approach has attempted to address, the exposures and outcomes considered in each study, and the contribution of each chapter to the triangulation exercise presented here.

\blandscape
<!----------------------------------------------------------------------------->
(ref:thesisOverview-caption) Summary of studies included in this thesis, and used as evidence sources in this triangulation exercise. Note, Chapter \@ref(sys-rev-tools-heading) is intentionally not included in this table, as it describes a research tool rather than a research study.

@@

As discussed in the introduction to this Chapter (Section \@ref(triangulation-overview)), as part of a narrative synthesis of the evidence all information sources were grouped by outcome, and systematically compared and contrasted. Potential reasons for heterogeneity between study designs were examined with specific reference to the risk-of-bias assessments performed.

@@

In addition to the qualitative discussion of evidence, I attempted to integrate the numerical results of the multiple approaches. This proof-of-concept approach incorporates recent advancements in the way that bias in results is assessed (move from checklists to domain-based assessment tools)[@sterne2019] and existing methods for bias-/indirectness-adjusted meta-analysis[@turner2009] to to illustrate how causal questions could be addressed under this quantitative triangulation framework.

This proposed framework involves several steps, as described in detail in the following sections. In summary these steps are:

1. Define the causal question of interest
1. Identify of relevant evidence sources and standardisation of effect directions
1. Assess and visualise the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Caclulate bias-/indirectness-adjusted results and perform meta-analysis

In order to illustrate the bias-/indirectness adjustment process detailed in the subsequent sections, the process for calculating the adjusted estimate for a single study is described in detail for a single study. <!-- TODO CITATION NEEDED need to cite single paper here --> Following this, the process for an example causal question as a case study, as defined below.

@@

Given that this quantitative analysis is intended as a proof-of-concept for the framework, rather than aiming to produce findings that should guide clinical practice, I consider only one of the multiple possible causal questions related to the broad topic of the effect of lipids on dementia outcomes.

Using the ROBINS-E framework I defined the parameters of my causal question of interest:

*	_Population of interest_: General population
*	_Exposure of interest_: Low density lipoprotein cholesterol
*	_Exposure window of interest_: Mid-life (45-60)
* _Summary measure of exposure over time_: Average exposure

This question was chosen for a number of reason. Firstly, given the long prodomal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period is particularly important. Secondly, examining the causal impact of lipids on a specific outcome (in this case, Alzheimer's disease), rather than a composite outcome (all-cause dementia), focuses on an single causal pathway rather than including conditions that likely have very different mechanisms of disease. Finally, and relevant to its use as a case study for the quantitative triangulation approach, this question had several relevant sources of evidence available (versus, for example, any question looking at vascular dementia).

@@

Once the causal question of interest had been defined, relevant individual results were obtained from the data sources described in Section \@ref(tri-data-sources). In a broad sense, this meant pulling out studies that examined the relationship between LDL-c and Alzheimer's disease risk, either directly (non-randomised studies of LDL-c levels) or indirectly (non-randomised studies of statins, Mendelian randomisation studies using genetic instruments for LDL-c levels).

Once the set of relevant results were identified, effects were standardised to refer to the risk of Alzheimer's disease resulting from a "reduction" in LDL-c. For studies of statins, th. However, for both non-randomised studies and Mendelian randomisation analysis of lipid levels, the effect is usually reported per 1-SD increase in LDL-c. In this case, the effect estimates were inverted to ensure consistency across study designs.

@@

The risk of bias tools used to assess each result are described in detail in Section \@ref(risk-of-bias). In summary, RTCS were assessed using the RoB 2 tool;[@sterne2019] NRSI  using the ROBINS-I tool;[@sterne2016] NRSE using the ROBINS-E tool (see Section \@ref(#rob-tools-nrse));[@morganr2020; @french2019] and MR studies using the Mamluk et al. tool.[@mamluk2020]

For the sake of consistency, I mapped the different acceptable judgements in each risk-of-bias tool to a harmonised set of judgements: "Low", "Moderate","High". The exact mapping can be seen in Table \@ref(tab:robLevelsMapping-table). Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis.[@sterne2016]

@@

In addition to assessing the extent of bias in each domain, as defined by the three levels detailed above, I attempted to record the predicted direction and type of bias. This was achieved via an additional question at the end of each bias domain, which asked assessors to predict the expected direction of bias in that domain. Note that this question is currently only formally included in two existing tools (ROB2/ROBINS-I), but I employed an identical approach during assessment of non-randomised studies of exposure and Mendelian randomisation studies. Acceptable responses to this question included:

* "Favours experimental"/"Favours comparator" (defines an additive bias)
* "Towards null"/"Away from null" (defines a proportional bias)
* "Unpredictable"

As indicated in the options above, the response to this question is also used to determine the predicted type of bias (additive or proportional - see Section \@ref(tri-terminology)) in addition to the direction. For additive biases, determining the absolute direction of bias was simply a case of whether the bias shift the effect estimate to the left or right on an imaginary forest plot. For example, if a bias was predicted to favour the intervention, then the estimate would be shifted to the right.

In contrast, for proportional biases, whether the estimate should be increased or decreased proportionally depends on the current position of the point estimate relative to the null. For example, if the effect estimate represents a protective effect (below the null), then bias towards the null would be adjusted for by moving the effect estimate propotionally to the right. In contrast, if the effect of the intervention is harmful (effect estimate above the null), bias towards the null would be adjusted for by moving the effect estimate propotionally to the left.

<!----------------------------------------------------------------------------->
(ref:exampleDirection-cap) Illustration of calculating the absolute direction of biases (indicated by arrows), based on bias type. For additive biases, the direction is consistent regardless of the position of the effect estimate relative to the null. In contrast, the position of the effect estimate is accounted for when determining the absolute direction of proportional biases.

(ref:exampleDirection-scap) Illustration of calculating the absolute direction of biases, based on bias type.

```{r exampleDirection, include=F}
```

```{r exampleDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:exampleDirection-cap)', out.width='100%', fig.scap='(ref:exampleDirection-scap)'}
knitr::include_graphics(file.path("figures/tri/exampleDirection.png"))
```
<!----------------------------------------------------------------------------->

Note that for some domains, only one type of bias is allowed. For example, in the confounding domains in the ROBINS-I/ROBINS-E tools, only the "Favours experimental"/"Favours comparator", and "Unpredictable" options are available, meaning that bias in this domain will always be considered additive. Additionally, the predicted direction and type of bias were not recorded for domains at "Low" risk of bias.

@@

To aid with the triangulation exercise and to act as a sense-check, I created a new method to visualise the extent, direction and type of bias in each result to enable detailed comparison across different studies contributing to a synthesis. Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), these "bias direction" plots show the  level of bias across the domains in each study is reported using coloured blocks. However, in contrast to the previous plots, symbols are used to illustrate the predicted absolute direction of bias in that domain, as discussed in the previous section. Distinct symbol pairs are used to denote additive and proportional biases.

Convience functions in the `triangulate` package (see Section \@ref(tri-software)) were used to preprocess risk of bias data for use with both bias-direction plots and adjusted meta-analyses, as discussed in subsequent sections. These bias direction plots themselves were created using the risk-of-bias visualisation tool described in Appendix \ref(?).

@@

####	Assign modifying values to risk/direction of bias
<!-- TODO Note that they are on the log scale - Lines 45-50 of addbias -->

Following definition of the extent, type and direction of bias in each domain, I assigned a prior distribution on the log scale to each combination. For example, a "high" additive bias in $j$th domain of the $i$th study is deifined as $\delta_{i,j}^{\mathrm{High}}$ and uncertainty around this estimate is given by a distribution:

@@

The sign of $\mu_{i,j}^{\mathrm{High}}$ is defined by the absolute direction of the bias. If the bias is expected to pull the effect to the left, the position is given a negative sign, and if it is expected to pull the effect to the right, it is given a positive sign.

One key feature of the domain-based risk-of-bias domains is that the domains are considered interchangeable - i.e. a "high" judgement in one domain is equivalent to a "high" judgement in any other. As such, for this analysis, I assumed that the distributions assigned to each level of bias are identical across all values of $j$, that is, are identical for all domains in the risk of bias assessment tool. To illustrate, for additive biases these means:

@@

Reasonable values for the position and spread of the prior distributions were informed by a previous reported expert bias elicitation exercise.[@turner2009] Using open data from that study, I calculated the mean and maximum values for the mean and variance of the adjustment distributions across all biases considered by that exercise (see Appendix \@ref(?) for more details on how these were calculated). For the purposes of this illustrative analysis and to highlight the generalised nature of the code used to perform the analysis (see Section \@ref(tri-software)), I used the calculated values to inform two sets of modifying distributions (Table \@ref(tab:priorsAdd-table)) for additive biases, and a single set for proportional biases. In the first scenario, the relationship between the position assigned to each additive judgement is linear. In the second scenario, the step from "moderate" to high" is twice that from "low" to "moderate". Across both additive and proportional biases, the variance of the "high" risk--of-bias judgement distribution was defined to be greater than that for the "moderate" risk-of-bias judgement.

@@

(ref:priorsAdd-caption) Prior distributions for extent and type of bias under two scenarios, on the log scale. Note that the distributions for proportional biases were kept consistent across the two scenarios.

@@

 <!-- TODO Link this to open data! -->

Where the direction of bias in a domain was unpredictable for a given result, the mean of the adjustment distribution for that domain was set to 0 but the appropriate variance for the recorded extent of bias was retained. In other words, adjusting for bias in a domain with an unpredictable direction of bias has no impact on the position of the effect estimate, but does increase uncertainty around it.

@@

An identical approach was employed to assess and assign prior distributions to the risk of indirectness in each result. The target question for each result was compared against the causal question of interest with respect to three domains (population, intervention/exposure, outcome), <!-- TODO CITATION NEEDED --> again using the scale of "Low/"Moderate"/"High" to quantify the extent of indirectness.

A similar approach was used to define prior distributions for external bias or indirectness. I used a similar enforced scale of "Low/"Moderate"/"High" indirectness versus each of the 4 aspects of the target causal question. (recap). All indirectness were defined _a priori_ as being proportional in nature, in line with previous work on this topic.[@turner2009]

As crude illustrative example, consider the causal question of the interest in this exercise, the effect of LDL-c in mid-life on Alzheimer's disease. Here, studies of lipids levels in this time period would require minimal adjustment, whereas studies examining statin use (indirect exposure/intervention) in late-life (indirect population) would be downweighted due to reduced relevance to the causal question.

@@

As per the method reported in, total additive and proportional bias and indirectness were calculated for each result. Briefly,

<!-- TODO Talk about how total add/prop biases calculated -->

Once the total additive and proportional bias and indirectness for each result were computed, they were used to define an adjusted estimate for each result included in the synthesis.

@@

The standard error of this estimate is then calculated as:

@@

Once each individual result has been adjusted, I performed a random effects meta-analysis using the unadjusted and adjusted (under the two scenarios defined above) results. <!-- TODO CITATION NEEDED -->

@@

Total additive bias (on the log scale) was `r hold()`. Similarly, total additive indirectness was (),

@@

`r hold()`

The unadjusted effect for this study was `r hold()` and the adjusted effect estimate for this study was `r hold()`

@@

#### Case study: effect of LDL-c at midlife on Alzheimer's disease

From the data sources identified in Section \@(), I identified X results relevant to my causal question of interest. <!-- TODO CITATION NEEDED will need to cite these! --> The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:ldlAdBiasDirection), stratified by study design.

@@

(ref:ldlAdBiasDirection-cap) Bias direction plot, summarising the internal biases related to effect of LDL-c on Alzheimer's disease, stratified by study design.

@@

```{r ldlAdBiasDirection, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:ldlAdBiasDirection-cap)', out.width='100%', fig.scap='(ref:ldlAdBiasDirection-scap)'}
knitr::include_graphics(file.path("figures/tri/midlife_AD.png"))
```
<!-- TODO Bias direction plot is not producing subgroup test for some reason -->
&nbsp;<!----------------------------------------------------------------------->

The results of the unadjusted and the bias-/indirectness-adjusted (under the Scenario 1 of additive biases) are shown in Figure \@ref(fig:fpLdlAd). I observed no substantial change to the overall effect (unadjusted `r unadj_effect`; adjusted `r adj_effect`), though adjustment for bias and indirectness substantially reduced the heterogeneity between studies (unadjusted $\tau^2$ = `r unadj_tau2`, $I^2$ = `r unadj_I2`; adjusted  $\tau^2$ = `r adj_tau2`, $I^2$ = `r adj_I2`). However, as can been seen from the right panel in the figure, this reduction in heterogeneity between studies is largely achieved via a reduction in precision.

To explore the different impacts of the two defined scenarios for additive bias, I plotted these:

<!-- TODO Finish -->

@@

<!-- TODO This should really be stratified by type -->
<!-- TODO Standardise effect estimates to same direction -->
<!-- TODO Need to add my own evidence -->

(ref:fpLdlAd-cap) Random effects meta-analysis using unadjusted and bias-/indirectness-adjusted results.

(ref:fpLdlAd-scap) Random effects meta-analysis using unadjusted and bias-/indirectness-adjusted results.

```{r fpLdlAd, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:fpLdlAd-cap)', out.width='100%', fig.scap='(ref:fpLdlAd-scap)'}
knitr::include_graphics(file.path("figures/tri/fp_paired_midlife_ldl_ad.png"))
```

@@

<!-- IDEA Do turner et al standardise effect direction? -->

@@

This chapter has narratively synthesised the existing and new evidence identified and produced by the previous chapters, highlighting the absence of any consistent association between any blood lipid and any dementia outcome in the context of the bias in each approach.

In addition, it has sought to provide a generalised framework for quantitative triangulation, building on existing methods for systematic domain-based risk of bias assessment and bias-/indirectness-adjusted meta-analysis. While a lack of information on the prior distributions of bias and indirectness in non-randomised studies precluded, to illustrate the method, I considered the example causal question of the effect of mid-life LDL-c on Alzheimer's disease risk derived from a previous expert elicitation exercise.

@@

###	Limitations

@@

#### Defining reasonable prior distributions for bias and indirectness

The key limitation of this quantitative synthesis presented in this chapter is the ambiguity

It is the reason that only one causal question is addressed as a case study, as to base clinical recommendations on a method dependent on non-empirical values would be inappropriate.

@@

New software for performing risk of bias assessments, being developed by the Bristol Appraisal and Review of Research group has this as a secondary methodological aim.

@@

@@

Similar there are several examples of usmeta to define priors for other statistical terms in meta-analysis. For example, previous work has looked at defining prior distributions of heterogeneity across different topic areas using a large collection of meta-analyses from the Cochrane Collaboration.<!-- TODO Cite turner etc -->

<!-- TODO Talk about the fact recent paper does not favour bias-adjustment -->

@@

#### Low and critical risk of bias

@@

### Strenghts

In addition, the generalised framework will allow for future researchers to apply (see Section \@ref())

Appreciation that both the risk of bias and indirectness of a result should be assessed is growing. Indeed, some existing risk of bias for the assessment of diagnostic test accurary[@whiting2011a] and prediction models[@moons2019] consider indirectness (termed applicability in these tools) alongside the assessment of bias. <!-- TODO Cite PROBAS and QUADAS --> <!-- TODO CITATION NEEDED cite paper that shows that ROB is not being intergrated into synthesis --> The approach proposed here will capitalise on this, providing users of domain based tools with a further avenue by whih to incorporate the results of their risk-of-bias assessments quantitatively.

Additional if users are interested in only a summary of the evidence rather than assessing results in relation to a target qusetion, then internal biases can be adjusted for on their own.

@@

### Future research

As the need and appetite for synthesis of different sources of evidence (trials, non-randomised studies, etc.), future iterations of the risk of bias tools should take this into account. Simple steps like harmonising the risk of bias levels across tools would be a good first move towards a more integrate suite of tools. Similarly, software that implements the risk-of-bias tools should both allow for direction of bias question and carefully consider how this data will be exported.

In the mean time, the extent of the impact of knowledge of the results could be analysed using existing data. For example, a re-analysis of the data presented in a recent paper where both expert elicitation and structured

### Triangulation as an extension of evidence synthesis

@@

### Outputs {#tri-software}
